---
title: "Class 10"
author: "Aaron Liu A13908620"
format: html
---

##Comparative structure analysis of Adenylate Kinase

We will use the `bio3` package
We will also use the `mas` package from bioconductor
First we need to install the `BiocManager` using `install.packages()`

```{r}
library(bio3d)

```

We use `BiocManager::install()` to install `msa`
First we will pick a sequence of interest
```{r}
aa<-get.seq("1ake_A")
aa
```
Search for all related structure

```{r}
#Please not run this again
#b<-blast.pdb(aa)
hits<-plot(b)
```

```{r}
hits$pdb.id
```
>Side notes: save the results so we don't run blast again. 

```{r}
save(hits, b, file="blast_results.Rds")
```

```{r}
load("blast_results.Rds")
hits
```

Download all these related structures from the database with`get.pdb()`

```{r}
library(BiocManager)
library(bio3d)
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
Align and superpose structures
```{r}
pdbs<-pdbaln(files, fit=T, exefile="msa")
```

```{r}
pdbs
```
##PCA

We can do a PCA on the coordinate data i.e the `$xyz` values
```{r}
pc.xray<-pca(pdbs)
plot(pc.xray)

```
```{r}
dim(pdbs$xyz)
```

```{r}
mktrj(pc.xray,file="pca_results.pdb")
```



