---
title: "Class11: alphafold 2 analysis"
author: "Aaron Liu A13908620"
format: pdf
---
```{r}
library(bio3d)
```

Here we post process and insepct our modeling from AlphFold2


My results from AF live in the folder/directory `hivprdimer_23119`

```{r}
results_dir<-"hivprdimer_23119/"
pdb_files<-list.files(results_dir, pattern=".pdb", full.names = T)
```

We first need to align and supperpose the PDB models and wecan use the `pdbaln` function for this:

```{r}
pdbs<-pdbaln(pdb_files, fit=T, exefile='msa')
```
##The RMSD matrix

A common measure of structural dis-similarity is called RMSD

```{r}
rd<-rmsd(pdbs)
rd
```
```{r}
library(pheatmap)
rownames(rd)<-paste0("m",1:5)
colnames(rd)<-paste0("m",1:5)
pheatmap(rd)

```
Let's view these in Mol*. Here we want the fitted coords
```{r}
xyz<-pdbfit(pdbs, outpath="fitted")
```

Focus on the most invariant part so the fitting and supperposiiton will fit well. 

```{r}
core<-core.find(pdbs)
```

```{r}
core.inds<-core
xyz<-pdbfit(pdbs, inds=core.inds, outpath="core_fitted")
```

To evaluate how good multi-chain or multi-domain models are we need to look at the PAE scores (predicted aligned error)


They are output in the json file 
```{r}
pae_files<-list.files(results_dir, pattern="000.json", full.names=T)
pae_files
```
```{r}
library(jsonlite)
pae1<-read_json(pae_files[1], simplifyVector=T)
attributes(pae1)
pae1$max_pae
```

```{r}
pae5<-read_json(pae_files[5], simplifyVector=T)
pae5$max_pae
```
```{r}
plot.dmat(pae5$pae, 
          xlab="Residue No.",
          ylab="Residue No.", 
          zlim=c(0,30))
```


```{r}
plot.dmat(pae1$pae, 
          xlab="Residue No.",
          ylab="Residue No.", 
            zlim=c(0,30))

```


##Main point

We can run Alpha fold in google compute 
We can read these results into R 



